Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.

### 2019-02-28 papers/CADE19/paper.tex
fix some todos, add combined converge/decay results

\documentclass{llncs}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage[dvipsnames]{xcolor}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}
%\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage~\newpage
\section{Optimising Age/Weight Ratios} \label{sec:awr:study}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of active clauses reported by Vampire after successful 1-second runs on a TPTP problem.
In between the peaks on either side, the function of \(L\) is discontinuous with large peaks and troughs, but follows an approximate trend and settles toward the global optimum.
This is typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

\subsection{Experiments}
%Changing Vampire's fixed AWR has a significant effect on the number of clauses required to be processed before a proof is found.
As an initial illustrative example of how varying the AWR effects the number of clauses required to be processed before a proof is found consider Figure \ref{fig:example-optimal-awr}. This shows \todo{FINISH}

  shows the effect of varying AWR on an illustrative problem from TPTP: a smaller number of active clauses means that fewer clauses were processed, which in general means that a proof was found faster\footnote{
It should be noted that if a small number of clauses are extremely expensive to process it may be slower than a larger number of less-expensive clauses, but in general this is a good heuristic measure for prover performance.
It also avoids reproducibility issues involved with using system timing approaches.
}.
On this problem, a good AWR value is over 400\% better by this metric than the worst AWR value.

This experiment was repeated on the whole TPTP problem set, excluding problems Vampire does not support, problems which were not solved by the default mode in under a second for any AWR value, and problems solved in pre-processing.
The whole set yielded similar AWR data for 7,947 problems.
These data show that while there is no ``best'' AWR for the whole set (Table \ref{tab:no-best-awr}), choosing a good AWR value for a problem is well-rewarded (Figure \ref{fig:optimal-awr-improvement}).

\noindent
 \todo{Can we also capture the solved/unsolved balance? Perhaps just say what \% of solved problems were solved using that ratio.}
\todo{This discussion should be expanded. Results need more explanation and we should summarise e.g. yes we confirmed assumptions.}\\
\todo{Can we extract some general information on correlation (or lack thereof) between best AWRs and syntactic features?}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{optimal-awr-improvement}
	\caption{The distribution of the relative performance gain choosing the best AWR value creates, relative to the worst AWR value.}
	\label{fig:optimal-awr-improvement}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variable AWR for Vampire}

@@

\todo{A little more explanation of the Fig... eg what is AWR trend?}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{random-walk}
	\caption{The AWR series that produced the lowest number of activations on a particular problem, smoothed in order to show the actual effect on proof search. This is a search strategy that a single fixed AWR cannot reproduce.}
	\label{fig:random-walk}
\end{figure}

@@

\todo{Given that we don't, perhaps we should change the parameter to \emph{rate} or something else?}

@@

Together, these new settings allow a number of new option combinations, which can be used in conjunction with Vampire's portfolio \emph{CASC mode} pending integration into the strategy schedules.
This version of the prover is currently in a separate branch in the main Vampire repository\footnote{\texttt{https://github.com/vprover/vampire/tree/awr-shapes}}.

@@

\subsection{Benchmark Performance}
Figure \ref{fig:results} shows the performance in terms of solved problems of all the configurations tested.
These data show that configurations which are more similar to the baseline (i.e. slow decay or fast convergence) achieve more similar performance, as expected.
In total, 134 problems were solved by the new configurations that were not solved by the baseline.

\begin{figure}
\centering
\begin{tabular}{l l l l l l}
	Configuration & Frequency & Proved & Fresh & Uniques & \(u\)-score\\
	\hline
	baseline & -- & 13057 & 0 & 1 & 714.2\\
	converge & 1 & 13039 & 24 & 3 & 714.3\\
	converge & 5 & 13029 & 27 & 1 & 709.5\\
	converge & 10 & 13028 & 35 & 5 & 714.3\\
	converge & 50 & 13015 & 45 & 5 & 712.8\\
	converge & 100 & 12976 & 51 & 1 & 705.9\\
	converge & 500 & 12895 & 63 & 4 & 698.3\\
	converge & 1000 & 12837 & 52 & 0 & 688.6\\
	converge & 5000 & 12775 & 53 & 1 & 682.4\\
	converge & 10000 & 12751 & 53 & 0 & 678.7\\
	decay & 1 & 12698 & 48 & 1 & 673.6\\
	decay & 5 & 12702 & 51 & 1 & 674.9\\
	decay & 10 & 12698 & 48 & 1 & 674.2\\
	decay & 50 & 12712 & 49 & 2 & 679.1\\
	decay & 100 & 12726 & 46 & 1 & 678.8\\
	decay & 500 & 12795 & 29 & 1 & 685.5\\
	decay & 1000 & 12860 & 29 & 2 & 692.6\\
	decay & 5000 & 12982 & 16 & 2 & 707.1\\
	decay & 10000 & 13002 & 7 & 0 & 706.3\\
\end{tabular}
\caption{
Results for the tested configurations.
\emph{Proved} refers to the total number of problems a configuration solved.
\emph{Fresh} is the number of problems a configuration solved which were not solved by the baseline.
\emph{Uniques} is the number of problems a configuration solved which were not solved by any other configuration.
\emph{\(u\)-score} is a refined unique score which correlates to a configuration's utility in solving new problems, as used in Hoder \emph{et al.}~\cite{u-score}.
}
\label{fig:results}
\end{figure}

>>>>

\documentclass{llncs}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}
\usepackage[dvipsnames]{xcolor}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Optimising Age/Weight Ratios} \label{sec:awr:study}

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of activations reported by Vampire after successful 1-second runs on a TPTP problem.
In between the peaks on either side, the function of \(L\) is discontinuous with large peaks and troughs, but follows an approximate trend and settles toward the global optimum.
This is typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

\subsection{Experiments}
As an initial illustrative example of how varying the AWR effects the number of clauses required to be processed before a proof is found consider Figure \ref{fig:example-optimal-awr}. This shows \todo{FINISH}

  shows the effect of varying AWR on an illustrative problem from TPTP: a smaller number of activations means that fewer clauses were processed, which in general means that a proof was found faster\footnote{
It should be noted that if a small number of clauses are extremely expensive to process it may be slower than a larger number of less-expensive clauses, but in general this is a good heuristic measure for prover performance.
It also avoids reproducibility issues involved with using system timing approaches.
}.
On this problem, a good AWR value is over 400\% better by this metric than the worst AWR value.

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of activations reported by Vampire after successful 1-second runs on a TPTP problem.
In between the peaks on either side, the function of \(L\) is discontinuous with large peaks and troughs, but does follow an approximate trend and settles toward the global optimum.
These properties are typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

This experiment was repeated on the whole TPTP problem set, excluding problems Vampire does not support, problems which were not solved by the default mode in under a second for any AWR value, and problems solved in pre-processing.
The whole set yielded similar AWR data for 7,947 problems.
These data show that while there is no ``best'' AWR for the whole set (Table \ref{tab:no-best-awr}), choosing a good AWR value for a problem is well-rewarded (Table \ref{tab:awr-improvement}).

\noindent
 \todo{Can we also capture the solved/unsolved balance? Perhaps just say what \% of solved problems were solved using that ratio.}
\todo{This discussion should be expanded. Results need more explanation and we should summarise e.g. yes we confirmed assumptions.}\\

@@

\begin{table}
	\caption{Relative performance gain, showing the ratio in activations between the best AWR setting for a given problem and another base setting. We compare the base settings of 1:1 (Vampire's default), 1:5 (the best-behaved from Table \ref{tab:no-best-awr}), and the worst setting for the problem.}
	\centering
	\begin{tabular}{c | c | c c}
		& Maximum Gain & Mean Gain & (Standard Deviation)\\
		\hline
		1:1 & 133.56 & 1.26 & 1.63 \\
		1:5 & 133.67 & 1.44 & 1.70 \\
		(worst) & 222.01 & 3.95 & 7.60 \\
	\end{tabular}
	\label{tab:awr-improvement}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variable AWR for Vampire}

@@

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{random-walk}
	\caption{The AWR series that produced the lowest number of activations on a particular problem, smoothed in order to show the actual effect on proof search. This is a search strategy that a single fixed AWR cannot reproduce.}
	\label{fig:random-walk}
\end{figure}

@@

@@

Together, these new settings allow a number of new option combinations, which can be used in conjunction with Vampire's portfolio mode pending integration into the strategy schedules.
This version of the prover is currently in a separate branch in the main Vampire source repository\footnote{\texttt{https://github.com/vprover/vampire/tree/awr-shapes}}.

@@

\subsection{Benchmark Performance}
Figure \ref{tab:results} shows the performance in terms of solved problems of all the configurations tested.
These data show that configurations which are more similar to the baseline (i.e. slow decay or fast convergence) achieve more similar performance, as expected.
In total, 134 problems were solved by the new configurations that were not solved by the baseline.

\begin{table}
	\caption{
Results for the tested configurations.
\emph{Proved} refers to the total number of problems a configuration solved.
\emph{Fresh} is the number of problems a configuration solved which were not solved by the baseline.
\emph{Uniques} is the number of problems a configuration solved which were not solved by any other configuration.
\emph{\(u\)-score} is a refined unique score which correlates to a configuration's utility in solving new problems, as used in Hoder \emph{et al.}~\cite{u-score}.
	}
	\centering
	\begin{tabular}{l l l l l l}
		Configuration & Frequency & Proved & Fresh & Uniques & \(u\)-score\\
		\hline
		baseline & -- & 13057 & 0 & 1 & 714.2\\
		converge & 1 & 13039 & 24 & 3 & 714.3\\
		converge & 5 & 13029 & 27 & 1 & 709.5\\
		converge & 10 & 13028 & 35 & 5 & 714.3\\
		converge & 50 & 13015 & 45 & 5 & 712.8\\
		converge & 100 & 12976 & 51 & 1 & 705.9\\
		converge & 500 & 12895 & 63 & 4 & 698.3\\
		converge & 1000 & 12837 & 52 & 0 & 688.6\\
		converge & 5000 & 12775 & 53 & 1 & 682.4\\
		converge & 10000 & 12751 & 53 & 0 & 678.7\\
		decay & 1 & 12698 & 48 & 1 & 673.6\\
		decay & 5 & 12702 & 51 & 1 & 674.9\\
		decay & 10 & 12698 & 48 & 1 & 674.2\\
		decay & 50 & 12712 & 49 & 2 & 679.1\\
		decay & 100 & 12726 & 46 & 1 & 678.8\\
		decay & 500 & 12795 & 29 & 1 & 685.5\\
		decay & 1000 & 12860 & 29 & 2 & 692.6\\
		decay & 5000 & 12982 & 16 & 2 & 707.1\\
		decay & 10000 & 13002 & 7 & 0 & 706.3\\
		\hline
		converge & (combined) & 13167 & 117 & -- & --\\
		decay & (combined) & 13106 & 93 & -- & --\\
	\end{tabular}
	\label{tab:results}
\end{table}
